---
title: "Midterm Note"
output: html_notebook
---
#This note have details related to NA(1-3), merging table(,Maps(4-), and Shiny apps()
#Replace Value to NAs(for more details: refer to labs 13.1)
1. if there is only one specific value (like -999) with NA across the entire data set./ is it is a string then
```{r}
life_history%>%
  replace_with_na_all(condition=~.x ==-999)%>%
  miss_var_summary() #mutate(newborn=na_if(newborn, "not measure"))
```

2. if there is mutiple values you want to replace with NA(use %in%)
```{r}
amniota_no_na<-amniota %>% 
  replace_with_na_all(condition=~.x %in% c(-999, -999.00, -999.00000, -999.000))
```

by selecting which variables
```{r}
amniota_no_na <- amniota_no_na %>%
  mutate(across(c(birth_or_hatching_weight_g, 
                  adult_body_mass_g,
                  no_sex_body_mass_g),
                ~na_if(., 0)))
```

3. count numbers of NA and its percentage, if also need the total number of na
```{r}
miss_var_summary(amniota_no_na)
```

```{r}
amniota_no_na%>%
  miss_var_summary()%>%
  summarize(total_na=sum(n_miss))
```

4.
