---
title: "Pivot Longer"
output: html_notebook
---

pivot_loner 9.1
pivot_wider&separate 9.2
summarize()` is a verb in `dplyr` that allows us to produce summaries of data. 8.1
group_by()` allows us to group data by a categorical variable(s) of interest.  8.1

#distinct() looks for all unique observations in rows 6.1
What if we want to know which islands each species occurs on?
```{r}
penguins %>%
  distinct(species, island,.keep_all=T) #kepp all kept all the variables
```

##if_else()
```{r}
mammals %>%
  select(genus, species, newborn) %>%
  mutate(newborn_new=ifelse(newborn == -999.00, NA, newborn)) #if,(action), else
```

1. Following the example above, convert all -999 values in the mammals dataframe to NA.
```{r}
mammals <- mammals %>% #this step is important
  mutate(across(c(mass, 
                  wean_mass, 
                  gestation, 
                  max_life, 
                  newborn, 
                  weaning, 
                  litter_size, 
                  afr, 
                  litters_year), 
                ~ifelse(.==-999, NA, .)))
```

## Counts 7.1
```{r}
msleep %>%
  count(vore) %>%
  ggplot(aes(x=vore, y=n))+geom_col() #use geom_col when we have x and y
```

#n_distinct() is a handy way of cleanly presenting the number of distinct observations.
Here we show the number of distinct genera over 100 in body weight.
```{r}
msleep%>%
  filter(bodywt >100) %>%
  summarize(n_genera=n_distinct(genus))
```