---
title: "quick note from hw"
output: html_notebook
---

1.count number of dataset of a column
```{r}
n_distinct(gapminder$country)
```


2. Which five countries have had the largest absolute population growth between 1925-2025? Show this as a table.
```{r}
gapminder %>%
  filter(year %in% c(1925, 2025))%>%
  select(country, year, pop)%>%
  pivot_wider(names_from = year, 
              values_from = pop, 
              names_prefix = "pop_")%>%
  group_by(country)%>%
  mutate(ppl_grwoth_absolute=pop_2025-pop_1925)%>%
  arrange(desc(ppl_grwoth_absolute))
```

3. If you want to draw the line for top five country the code will be like this
```{r}
gapminder %>%
  filter((year>=1925 &year<=2025)&country%in%c("India", "China", "Indonesia", "Pakistan", "USA"))%>%
  ggplot(aes(x=year, y=pop, color=country))+
  geom_point()+
  geom_line()+ 
  labs(title="Population Growth Over Time", x="Year", y="Population Growth Over Time")
```

4. Has global life expectancy changed between 1925 and 2025? Show the min, median, mean, and max for all countries (combined) in the dataset.
```{r}
gapminder %>%
  filter(year %in% c(1925, 2025))%>%
  group_by(year)%>%
  summarize(min_lex=min(lex, na.rm=T),
         max_lex=max(lex, na.rm=T),
         mean_lex=mean(lex, na.rm=T),
         median_lex=median(lex, na.rm=T))
```

5.Make a plot that shows the distribution of life expectancy for all countries in 1925 and 2025 (try using `geom_density`). What do you notice about the shift in life expectancy over time?
```{r}
gapminder %>%
  filter(year %in% c(1925, 2025))%>% #this give only 1925 and 2025
  mutate(year=factor(year))%>%
  ggplot(aes(x=lex, fill=year, color=year))+
  geom_density(alpha=0.3)+
  labs(title="Distribution of Life Expetancy, 1925 ad 2025",
       x="Life Expectancy(years)",
       y=NULL, 
       color="year",
       fill="year")
```

6. (hw10) Are there any "hotspots" for shark incidents in California? Show the top five counties with the most incidents.
```{r}
sharks%>%
  select(county, incident_num)%>%
  group_by(county)%>%
  count(incident_num) %>%
  summarize(total_incident=sum(n))%>%
  arrange(desc(total_incident))%>%
  slice_max(total_incident, n=5)
```

7. Are there months of the year when incidents are more likely to occur? Make a plot that shows the total number of incidents by month. Which month has the highest number of incidents? 
```{r}
sharks%>%
  select(month, incident_num)%>%
  group_by(month)%>%
  count(incident_num)%>%
  ggplot(aes(x=as.factor(month), y=n))+geom_col()+labs(title="Total number of incidents for each month", x="month", y="count of incident")
```

7. In the data, `mode` refers to a type of activity. Which activity is associated with the highest number of incidents?
```{r}
sharks%>%
  select(mode, incident_num)%>%
  group_by(mode)%>%
  count(incident_num, name="total_incident_mode")%>%
  summarize(total_incident_mode=sum(total_incident_mode, na.rm=T))%>%
  arrange(desc(total_incident_mode))
```