---
title: "Graphing stuff"
output: html_notebook
---

- fill is interchangable with group it is just group won't color the graph

#advanced graphing(11.1)
```{r}
options(scipen=999) #cancels the use of scientific notation for the session
```

#1. Line Graph
```{r}
gap %>%
  filter(year>=2000 & year <= 2025 & country=="USA") %>%
  ggplot(aes(x=year, y=gdp))+
  geom_line(color="red")+ #line graph
  geom_point(shape=5)+ #choose the shape
  scale_x_continuous(breaks=seq(2000,2025, by=5))+  
  #scale_x_continuous() modifies the x-axis for continuous numeric variables
  #breaks = specifies the exact tick marks to display on the x-axis.
  #seq(2000, 2025, by = 5) generates a numeric sequence from 2000 to 2025
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="GDP per Capita in the USA, 2000-2025", x="Year", y="GDP")
```

for function scale_x_continuous
```{r}
scale_x_continuous(
  breaks = 2000:2025,
  minor_breaks = NULL
  # breaks = 2000:2025 creates tick marks at every integer year
  # from 2000 to 2025 (2000, 2001, 2002, ..., 2025).
  #
  # minor_breaks = NULL removes the smaller secondary tick marks
  # that ggplot would normally add between major breaks.
)
```

#2. Histogram 
```{r}
gap%>%
  filter(year==1935)%>%
  ggplot(aes(x=lex))+
  geom_histogram(color="black", fill="turquoise4", bins=30)+
  labs(title="Distribution of Life Expectancy")
```

#3. Density Graph
```{r}
gap%>%
  filter(year==1935)%>%
  ggplot(aes(x=lex))+
  geom_density(color="black", fill="turquoise4", alpha=0.7)+ #alpha is the transparency
  labs(title="Distribution of Life Expectancy")
```

Density+histogram
```{r}
gap%>%
  filter(year==1940)%>%
  ggplot(aes(x=lex))+
  geom_histogram(aes(y=after_stat(density)), color="black", fill="turquoise4")+
  geom_density(color="red")
```

#4.geom_col(10.1)
```{r}
life_history %>%
  count(order, sort=T) %>% 
  ggplot(aes(x=order, y=n))+geom_col()+labs(title="observations for each order of mammals ")+coord_flip() #flips the x and y axis
```

- bar plot of the mean mass for each order?
```{r}
life_history %>%
  group_by(order)%>%
  summarize(mean_mass=mean(mass, na.rm=T)) %>% 
  ggplot(aes(x=order, y=mean_mass))+geom_col()+coord_flip()
```

- make the graph 排列 with 順序 using `reorder`.
```{r}
life_history %>%
  group_by(order)%>%
  summarize(mean_mass=mean(mass, na.rm=T)) %>% 
  ggplot(aes(x=reorder(order, mean_mass), y=mean_mass))+geom_col()+coord_flip()+scale_y_log10()
```

-We can adjust the plot further by specifying the size and face of the text using `theme()`. 
- `rel()` option changes the relative size of the title.
-`hjust` controls the title position.
```{r}
life_history %>%
  ggplot(aes(x=mass, y=litter_size))+
  geom_point(na.rm=T)+
  scale_x_log10()+
  geom_smooth(method=lm, se=F, na.rm=T)+
  labs(title="Mass vs. Litter Size", x="Mass(g)", y="Litter Size")+
  theme(plot.title=element_text(size=rel(1.5), hjust=.5))
```

bar plot samples:
```{r}
homerange%>%
  ggplot(aes(x=taxon, fill=trophic_guild))+geom_bar(position="dodge")+theme(axis.text.x=element_text(angle=60, hjust=0.85))+labs(title="Observation by Taxon", x="Taxonomic Group", y=NULL)
```

#case_when function
- which allows us to calculate a new variable from other variables.
- allows us to specify multiple conditions
```{r}
income_cat2000<- gap %>%
  filter(year==2000)%>%
  mutate(income_category=case_when(gdp<=1100 ~ "low",
                                   gdp>1100& gdp <=4500 ~ "low_middle",
                                   gdp>4500 & gdp<=13900 ~ "upper_middle",
                                   gdp>13900 ~ "high")) %>%
  select(country, gdp, income_category) %>%
  mutate(income_caegory=factor(income_category,
                               levels=c("low",
                                        "low_middle",
                                        "upper_middle",
                                        "high")))
```

## Faceting(11.2)
- allow us to make muti-panel plots for easy comparison

#1. boxplot examples - facet_wrap()
- makes a ribbon of panels by a specified categorical variable and allows you to control how you want them arranged.
```{r}
homerange %>% 
  ggplot(aes(x=taxon, y=log10_mass, fill=thermoregulation))+
  geom_boxplot()+
  facet_wrap(~trophic_guild, nrow=2, labeller=label_both)+ 
  #trophic_guild is a column, sort by carnivore/herbivore
  theme(axis.text.x=element_text(angel=60, hjust=1),
        legent.position="none",
        strip_text=element_text(size=12, face="bold"))
```

#2. facet_grid() allows control over the faceted variable; it can be arranged in rows or columns. rows~columns
```{r}
homerange %>% 
  ggplot(aes(x=taxon, y=log10_mass, fill=thermoregulation))+
  geom_boxplot()+
  facet_grid(thermoregulation~., labeller=label_both)+ 
  theme(axis.text.x=element_text(angel=60, hjust=1),
        legent.position="none",
        strip_text=element_text(size=12, face="bold"))
```

- will also allow the comparison of two categorical variables, just remember a~b where a is rows and b is columns. 
```{r}
homerange %>% 
  ggplot(aes(x=taxon, y=log10_mass, fill=thermoregulation))+
  geom_boxplot()+
  facet_grid(trophic_guild~thermoregulation, labeller=label_both)+
  theme(axis.text.x=element_text(angle=60, hjust=1),
        legend.position="none",
        strip.text=element_text(size=8, face="bold"))
```

