---
title: "Homework 11"
author: "CHUEH-JU KUO"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above. For any included plots, make sure they are clearly labeled. You are free to use any plot type that you feel best communicates the results of your analysis.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
``` 

```{r}
options(scipen = 999)
```

## Data
For this homework, we are going to use the [Gapminder](https://www.gapminder.org/data/) data on population (pop), GDP per capita (gdp), and life expectancy (lex). These data were downloaded and compiled using the `gapminder_compile.Rmd` script in the data folder.  

Let's focus our attention on data from the past 100 years, between 1925 and 2025.
```{r message=FALSE, warning=FALSE}
gapminder <- read_csv("data/gapminder.csv") %>% 
  clean_names() %>% 
  filter(between(year, 1925, 2025))
```

1. Use the function(s) of your choice to get an idea of the overall structure of the data, including its dimensions, column names, classes, etc. 
```{r}
glimpse(gapminder)
```
```{r}
summary(gapminder)
```

2. How many countries are represented in this dataset?
There are 195 countries.
```{r}
n_distinct(gapminder$country)
```

3. Which five countries have had the largest absolute population growth between 1925-2025? Show this as a table.
The top five country is India, China, Indonesia, Pakistan and USA
```{r}
gapminder %>%
  filter(year %in% c(1925, 2025))%>%
  select(country, year, pop)%>%
  pivot_wider(names_from = year, 
              values_from = pop, 
              names_prefix = "pop_")%>%
  group_by(country)%>%
  mutate(ppl_grwoth_absolute=pop_2025-pop_1925)%>%
  arrange(desc(ppl_grwoth_absolute))
```

4. Make a plot that shows population growth over time for the top country you found in question #3.
```{r}
gapminder %>%
  filter((year>=1925 &year<=2025)&country=="India")%>%
  ggplot(aes(x=year, y=pop, color=country))+
  geom_point()+
  geom_line()+ 
  labs(title="Population Growth Over Time", 
       x="Year", 
       y="Population")

#note for myself:
#if you want to draw the line for top five country the code will be like this
#gapminder %>%
#  filter((year>=1925 &year<=2025)&country%in%c("India", "China", "Indonesia", "Pakistan", "USA"))%>%
#  ggplot(aes(x=year, y=pop, color=country))+
#  geom_point()+
#  geom_line()+ 
#  labs(title="Population Growth Over Time", x="Year", y="Population Growth Over Time")
```

5. Has global life expectancy changed between 1925 and 2025? Show the min, median, mean, and max for all countries (combined) in the dataset.
```{r}
gapminder %>%
  filter(year %in% c(1925, 2025))%>%
  group_by(year)%>%
  summarize(min_lex=min(lex, na.rm=T),
         max_lex=max(lex, na.rm=T),
         mean_lex=mean(lex, na.rm=T),
         median_lex=median(lex, na.rm=T))
```

6. Make a plot that shows the distribution of life expectancy for all countries in 1925 and 2025 (try using `geom_density`). What do you notice about the shift in life expectancy over time?
Life expectancy of all country is shifted from number around 30 to 70 for 2025. This indicated the increment of life spans of human.
```{r}
gapminder %>%
  filter(year %in% c(1925, 2025))%>%
  mutate(year=factor(year))%>%
  ggplot(aes(x=lex, fill=year, color=year))+
  geom_density(alpha=0.3)+
  labs(title="Distribution of Life Expetancy, 1925 ad 2025",
       x="Life Expectancy(years)",
       y=NULL, 
       color="year",
       fill="year")
```

7. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer in 2025? Show this as a plot. (Suggestion! Remove Monaco as an extreme outlier.)
I think more money does help you live longer in 2025 since the graph have a positive trend line.
```{r}
gapminder %>%
  filter(year==2025& country != "Monaco")%>%
  mutate(capita_gdp=gdp/pop)%>%
  ggplot(aes(x=capita_gdp, y=lex))+
  geom_point()+geom_smooth(method = "lm", se = FALSE)+
  scale_x_log10()+
  labs(title="Relatiobship between per capita GDP and Life Expectancy",
       x="Per Capita GDP",
       y="Life Expectancy")
```

8. Which five countries have had the highest absolute GDP per capita growth over the past 50 years? Show this as a table. (Suggestion! Remove Monaco as an extreme outlier.)
The top five countries are San Marino, Palau, St. Kitts and Nevis, Tuvalu, and Nauru.
```{r}
gapminder %>%
  filter(year %in% c(1975, 2025) & country != "Monaco")%>%
  mutate(capita_gdp=gdp/pop)%>%
  select(country, year, capita_gdp) %>%
  pivot_wider(names_from = year, 
              values_from = capita_gdp, 
              names_prefix = "capita_gdp_")%>%
  group_by(country)%>%
  mutate(abs_capita_gdp=capita_gdp_2025-capita_gdp_1975)%>%
  arrange(desc(abs_capita_gdp))
```

9. How does per capita GDP growth compare between these same five countries over the past 50 years? Show this as a plot.
```{r}
gapminder %>%
  filter((year>=1975 &year<=2025)&country%in%c("San Marino", "Palau", "St. Kitts and Nevis", "Tuvalu", "Nauru"))%>%
  mutate(capita_gdp=gdp/pop)%>%
  ggplot(aes(x=year, y=capita_gdp, color=country))+
  geom_point()+
  geom_line()+ 
  labs(title="Per Capita GDP Growth(1975-2025)", x="Year", y="Per capita GDP")
```

10. Do one analysis of your choice that includes a table and plot as outputs.
How is the life expectancy  and GDP per capita of Germany change around 1945(end of world war 2) to 1965?
```{r}
gapminder %>%
  filter(between(year, 1945, 1965), country == "Germany") %>%
  mutate(capita_gdp = gdp/pop) %>%
  select(year, lex, capita_gdp) %>%
  pivot_longer(col=c(lex, capita_gdp), names_to = "type", values_to = "value")
```


```{r}
gapminder %>%
  filter(between(year, 1945, 1965), country == "Germany") %>%
  mutate(capita_gdp = gdp/pop) %>%
  select(year, lex, capita_gdp) %>%
  pivot_longer(cols=c(lex, capita_gdp), 
               names_to = "type", 
               values_to = "value") %>%
  ggplot(aes(x=year, y=value))+
  geom_line() +
  geom_point() +
  facet_wrap(~ type,scales = "free_y", ncol =2) + #free y means each graph have itw own y
  labs(title = "Life expectancy and GDP per capita of Germany(1945–1965)",
       x = "Year", 
       y = NULL)
```
## Knit and Upload
Please knit your work as an .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean! 